### 2021-04-28 CL/CoverLetter.tex
Initial Overleaf Import

%
% harvard-letter.tex - an example latex file to illustrate harvardletter.cls
%
% Copyright 2019, Ahmad Zareei (ahmad@seas.harvard.edu)
%

\documentclass[11pt]{harvardletter}
\usepackage{color}
\usepackage{soul}

\newcommand{\EM}[1]{\noindent \color{cyan} (EM: #1)\normalcolor}
\newcommand{\KB}[1]{{\color{red}{Katia: #1}}}
\newcommand{\del}[1]{\noindent \color{red} \st{#1}\normalcolor}

% \name{\includegraphics{./Signature.eps}} % used as signature, if no signature is specified

\signature{\vspace{-10mm}\includegraphics[height=0.4in]{Signature.eps} \\\textbf{Ariel Amir}\\ \textit{Thomas D. Cabot Associate Professor of Applied Mathematics and Applied Physics,} \\ \textit{School of Engineering and Applied Sciences,}\\ \textit{Harvard University} \\ \textit{Phone: (617) 495-5818}\\ \textit{Email: arielamir@seas.harvard.edu}}

\date{} % if no date specified, today's date is used

% \subject{This is the subject of the letter} % optional subject line

\begin{document}

\begin{letter}{}
  %              {Prof. ... .... \\ 
  %              Harvard University\\ 
  %              Cambridge, MA 02148}

\hfill{March 21, 2021}
\vspace{20mm}

% \begin{table}
%     \centering
    \begin{tabular}{l l }
        \small{Title:}  & \textbf{\small{Temporal Evolution of flow in Porous Materials: From Homogenization to Instability}} \vspace{2mm} \\ 
        \small{Authors:}  & \small{Ahmad Zareei$^1$, Pan Deng$^1$, and Ariel Amir$^{1,2}$* (*corresponding author)} \vspace{2mm} \\
        \small{Affiliations:} & \small{$^1$Harvard John A. Paulson School of Engineering and Applied Sciences, Harvard University} \\
        & \small{$^2$Kavli Institute, Harvard University}
    \end{tabular}
% \end{table}

\vspace{5mm}

\opening{Dear Editor,}

We are excited to submit our work, ``Temporal Evolution of flow in Porous Materials: From Homogenization to Instability'' for consideration as an article in the \emph{Physical Review Letters}. In this study, we demonstrate that ....

As such, our study suggests that

For reviewers, we suggest the following experts:

\textbf{Karen Alim},Max Planck Institute for Dynamics and Self-Organization, G\"ottingen, Germany, \texttt{k.alim@tum.de}

\vspace{5mm}

\underline{\textbf{\small{This material has not been published and is not under consideration for publication elsewhere.}}}

\vspace{5mm}

We thank you for your consideration and look forward to hearing from you.

\vspace{5mm}

\closing{Sincerely,}

% \encl{manuscript.pdf}
% \ps{Please see the enclosed file.}
% \cc{Dr. Ahmad Zareei}

\end{letter}

\end{document}